Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-17 ICML-submission/appendix-multiclass-perceptron.tex
a version within page limit

\begin{theorem}[Mistake upper bound]
\label{theorem:mutliclass-perceptron-mistake-upper-bound}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, let $K$ be a positive
integer, let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled
examples in $V \times \{1,2,\dots,K\}$ that are weakly linearly separable with margin
$\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
then \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes.
\end{theorem}

>>>>

\begin{theorem}[Mistake upper bound]
\label{theorem:mutliclass-perceptron-mistake-upper-bound}
\chicheng{Cite~\cite{Crammer-Singer-2003} here?}
Let $(V, \ip{\cdot}{\cdot})$ be an inner product space, let $K$ be a positive
integer, let $\gamma$ be a positive real number and let $R$ be a non-negative real
number. If $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ is a sequence of labeled
examples in $V \times \{1,2,\dots,K\}$ that are weakly linearly separable with margin
$\gamma$ and $\norm{x_1}, \norm{x_2}, \dots, \norm{x_T} \le R$
then \textsc{Multiclass Perceptron} algorithm makes at most $\lfloor
2(R/\gamma)^2 \rfloor$ mistakes.
\end{theorem}

### 2019-01-17 ICML-submission/from-weak-separability-to-strong-separability.tex
Fix Makefile and citation
a version within page limit

In this section, we consider case when the examples are weakly linearly
separable. Throughout this section, we assume without loss of generality that
all examples lie in the unit ball $\B(0,1) \subseteq \R^d$.\footnote{Instead of
working with feature vector $x_t$ we can work with normalized feature vectors
$\widehat{x}_t = \frac{x_t}{\norm{x_t}}$. It can be easily checked that if
$(x_1,y_1), (x_2,y_2), \dots, (x_T,y_T)$ are weakly linearly separable with
margin $\gamma$ and $\norm{x_t} \le R$ for all $t$, then the normalized examples
$(\widehat{x}_1,y_1), \widehat{x}_2,y_2), \dots, (\widehat{x}_T,y_T)$ are weakly
linearly separable by margin $\gamma/R$.} Note that
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
alone does not guarantee a finite mistake bound in this setting, as weak linear
separability does not imply strong linear separability.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(0,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
\begingroup
\allowdisplaybreaks
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(2K-5/2) 2^{K-1}} \\
& \ \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
\endgroup
Then, the sequence of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, $\norm{\phi(x_1)}_{\ell_2},
\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$.
\end{theorem}

Using this theorem we derive a mistake bound for
Algorithm~\ref{algorithm:kernelized} with kernel function $k(x,x') =
\ip{\phi(x)}{\phi(x')}_{\ell_2}$.

@@

This corollary follows direcly from
Theorems~\ref{theorem:kernelized-upper-bound}
and~\ref{theorem:margin-transformation}.

@@

\paragraph{Overview.} The idea behind construction and analysis of the mapping
$\phi$ is polynomial approximation. Specifically, we give $k$ polynomials $p_1$,
$\ldots$, $p_k$, such that:
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, p_{y_t}(x_t) \ge \frac{\gamma'}2 \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \leq -\frac{\gamma'}2 \; .
\end{gathered}
\end{align}
As each polynomial $p_i$ can be written as $\ip{c_i}{\phi(x)}$ for some $c_i \in
\ell_2$, it immediately shows that the set of examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ is strongly linear separable.

@@

Under the weak linear separability assumption, we note that all examples from
class $i$ lie in
\[ R_i^+ = \bigcap_{j \in \cbr{1,\ldots,K} \setminus \cbr{i}} \cbr{x: \ip{w_i - w_j}{x} \geq \gamma}, \] and all examples from the remaining classes lie in
\[ R_i^- = \bigcup_{j \in \cbr{1,\ldots,K} \setminus \cbr{i}} \cbr{x: \ip{w_i - w_j}{x} \leq  -\gamma}. \]
Therefore, to satisfy Equations~\eqref{equation:poly-pos-i}
and~\eqref{equation:poly-neg-i}, it suffices to construct $p_i$ such that:
\begin{align*}
x \in R_i^+ & \Rightarrow p_i(x) \ge \frac {\gamma'} 2 \\
x \in R_i^- & \Rightarrow p_i(x) \le - \frac {\gamma'} 2
\end{align*}

According to the well known Stone-Weierstrass theorem (see
e.g.~\citet[Section~10.10]{Davidson-Donsig-2010}), on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, the indicator function corresponding to the intersection of
$m=K-1$ halfspaces. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. The polynomial separates examples in
one class from examples in the other classes. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial. We
construct two different polynomials with different parameters. The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
\begin{align*}
p(x)
& = p(x_1, x_2, \dots, x_d) \\
& = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\end{align*}
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$'s are real coefficients. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
Following the terminology of~\cite{Klivans-Servedio-2008},
the \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as the
Euclidean norm of the vector of the coefficients of the polynomial.

\begin{theorem}[Polynomial approximation of intersection of halfspaces I]
\label{theorem:polynomial-approximation-1}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $p(x) \ge 1/2$ \\ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
\item $p(x) \le -1/2$ \\ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$
\item $\displaystyle \deg(p) = \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$
\item $\displaystyle \norm{p} \le \left( 188 \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil \right)^{\frac{1}{2} \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil}$
\end{enumerate}
\end{theorem}

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$.
Define
$$
r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(1/\gamma) \right \rceil \; .
$$
Then, there exists a multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $\displaystyle p(x) \ge \frac{1}{4} \cdot 2^{s(s+1)rm}$ \\
for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$

\item $\displaystyle p(x) \le - \frac{1}{4} \cdot 2^{s(s+1)rm}$ \\
for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$

\item $\deg(p) \le (2s+1) rm$
\item $\norm{p} \le (2m-1/2) 2^m \cdot \left(2^{2s} rm (4s+2)^2 \right)^{(s+1/2)rm}$
\end{enumerate}
\end{theorem}

@@

The following lemma expresses any multivariate polynomial in $\R^d$
as a linear function in $\ell_2$ and gives an upper bound its the norm.

@@

\begin{proof}
Note that the polynomial $p$ can be written as
$p(x) = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c'_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}$.
We define $c \in \ell_2$ using the multi-index notation as
$$
c_{\alpha_1, \alpha_2, \dots, \alpha_d}
= \frac{c'_{\alpha_1, \alpha_2, \dots, \alpha_d} 2^{(\alpha_1 + \alpha_2 + \dots + \alpha_d)/2}}{\sqrt{\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}}
$$
for all tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ such that $\alpha_1 + \alpha_2 + \dots + \alpha_d \le \deg(p)$.
Otherwise, we define $c_{\alpha_1, \alpha_2, \dots, \alpha_d} = 0$. Clearly,
$\ip{c}{\phi(x)}_{\ell_2} = p(x)$.

Since
\begin{align*}
|c_{\alpha_1, \alpha_2, \dots, \alpha_d}|
& \le 2^{(\alpha_1 + \alpha_2 + \dots + \alpha_d)/2} |c'_{\alpha_1, \alpha_2, \dots, \alpha_d}| \\
& \le 2^{\deg(p)/2} |c'_{\alpha_1, \alpha_2, \dots, \alpha_d}| \; ,
\end{align*}
we have
\begin{align*}
\norm{c}_{\ell_2}
& \le 2^{\deg(p)/2} \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} (c'_{\alpha_1, \alpha_2, \dots, \alpha_d})^2} \\
& = 2^{\deg(p)/2} \norm{p} \; . \qquad \qedhere
\end{align*}
\end{proof}

Using the lemma and the polynomial approximation theorems, we can prove the
mapping $\phi$ maps any a weakly separable data to a strongly separable data
set.

\begin{proof}[Proof of Theorem~\ref{theorem:margin-transformation}]
First osberve that for any $t=1,2,\dots,T$,
$$
\norm{\phi(x_t)}_{\ell_2} = k(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
$$

Since the examples $(x_1, y_1), (x_2, y_2), \dots, (x_T, y_T)$ are weakly
linearly separable with margin $\gamma$ there are vectors $w_1, w_2, \dots, w_K$
satisfying \eqref{equation:weak-linear-separability-1} anf
\eqref{equation:weak-linear-separability-2}.

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors $(w_i - w_j)/2$ for
$j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors have norm at most
$1$.
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
imply that there exists multivariate polynomials $p_i:\R^d \to \R$ and $q_i:\R^d
\to \R$ such that
\begin{align*}
\deg(p_i) & = \lceil \log_2(2K-2) \rceil \cdot \left\lceil \sqrt{\frac{2}{\gamma}} \right\rceil \; , \\
\deg(q_i) & = (2s+1) r(K-1) \; .
\end{align*}
Futhermore, for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$,
$q_i(x_t) \ge \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and if $y_t \neq i$ then
$p_i(x_t) \le -1/2$, $q_i(x) \le - \frac{1}{4} \cdot 2^{s(s+1)r(K-1)}$ and
\begin{align*}
\norm{p_i} & \le \left(188 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\norm{q_i} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
By \autoref{lemma:norm-bound} there exists $c_i, c_i' \in \ell_2$ such that
$\ip{c_i}{\phi(x)} = p_i(x)$ and $\ip{c_i'}{\phi(x)} = q_i(x)$ and
\begin{align*}
\norm{c_i}_{\ell_2}
& \le \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \\
\norm{c'_i}_{\ell_2} & \le (2K-5/2) 2^{K-1} \\
& \qquad \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{(s+1/2)r(K-1)} \; .
\end{align*}
Define vectors $u_i, u_i' \in \ell_2$ as
\begin{align*}
u_i & = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; . \\
u_i' & = \frac{c_i'  \cdot \left(2^{2s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)}}{\sqrt{K} (2K-5/2) 2^{K-1}} \; .
\end{align*}

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$ and
$\norm{u_1'}^2 + \norm{u_2'}^2 + \dots + \norm{u_K'}^2 \le 1$.
Futhermore, for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma_1$ and
$\ip{u'_{y_t}}{x_t} \ge \gamma_2$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma_1$ and $\ip{u'_j}{x_t} \le - \gamma_2$. In other words,
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_1$ and also strongly linearly
separable with margin $\gamma_2$. Therefore, the examples are strongly
linearly separable with margin $\gamma' = \max\{\gamma_1, \gamma_2\}$.
\end{proof}

>>>>

In this section, we consider case when the examples are weakly linearly
separable. Throughout this section, we assume without loss of generality that
all examples lie in the unit ball $\B(0,1) \subseteq \R^d$.\footnote{Instead of
working with feature vector $x_t$ we can work with normalized feature vectors
$\widehat{x}_t = \frac{x_t}{\norm{x_t}}$. It can be easily checked that if
$(x_1,y_1), (x_2,y_2), \dots, (x_T,y_T)$ are weakly linearly separable with
margin $\gamma$ and $\norm{x_t} \le R$ for all $t$, then the normalized examples
$(\widehat{x}_1,y_1), (\widehat{x}_2,y_2), \dots, (\widehat{x}_T,y_T)$ are weakly
linearly separable by margin $\gamma/R$.} Note that
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
alone does not guarantee a finite mistake bound in this setting, as weak linear
separability does not imply strong linear separability.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(0,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
\begingroup
\allowdisplaybreaks
\begin{align*}
r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(4K-5) 2^{K-1}} \\
& \ \cdot \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
\end{align*}
\endgroup
Then, the sequence of labeled examples transformed by $\phi$,
namely $(\phi(x_1), y_1), (\phi(x_2), y_2), \dots,
(\phi(x_T), y_T)$, is strongly linearly separable with margin $\gamma' =
\max\{\gamma_1, \gamma_2\}$. In addition, for all $t$ in $\cbr{1,\ldots,T}$,
$k(x_t,x_t) \leq 2$.
\end{theorem}
%$\norm{\phi(x_1)}_{\ell_2},
%\norm{\phi(x_2)}_{\ell_2}, \dots, \norm{\phi(x_T)}_{\ell_2} \le 2$

Using this theorem we derive a mistake bound for
Algorithm~\ref{algorithm:kernelized} with kernel function $k(x,x') =
\ip{\phi(x)}{\phi(x')}_{\ell_2}$, under the weak linear separability assumption.

@@

This corollary follows directly from
Theorems~\ref{theorem:kernelized-upper-bound}
and~\ref{theorem:margin-transformation}.

@@

\paragraph{Overview.} The idea behind the construction and analysis of the mapping
$\phi$ is polynomial approximation. Specifically, we give $k$ polynomials $p_1$,
$\ldots$, $p_k$, such that:
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, p_{y_t}(x_t) \ge \frac{\gamma'}2 \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t} \\
p_i(x_t) \leq -\frac{\gamma'}2 \; .
\end{gathered}
\end{align}
As each polynomial $p_i$ can be written as $\ip{c_i}{\phi(x)}$ for some $c_i \in
\ell_2$, it immediately shows that the set of examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ is strongly linear separable.

@@

Under the weak linear separability assumption, we note that all examples from
class $i$ lie in
\[ R_i^+ = \bigcap_{j \in \cbr{1,\ldots,K} \setminus \cbr{i}} \cbr{x: \norm{x} \le 1, \ip{w_i^* - w_j^*}{x} \geq \gamma}, \] and all examples from the remaining classes lie in
\[ R_i^- = \bigcup_{j \in \cbr{1,\ldots,K} \setminus \cbr{i}} \cbr{x: \norm{x} \le 1, \ip{w_i^* - w_j^*}{x} \leq  -\gamma}. \]
Therefore, to satisfy Equations~\eqref{equation:poly-pos-i}
and~\eqref{equation:poly-neg-i}, it suffices to construct $p_i$ such that:
\begin{align}
x \in R_i^+ & \Rightarrow p_i(x) \ge \frac {\gamma'} 2
\label{eqn:r-plus}
\\
x \in R_i^- & \Rightarrow p_i(x) \le - \frac {\gamma'} 2
\label{eqn:r-minus}
\end{align}

According to the well known Stone-Weierstrass theorem~\citep[see
e.g.][Section~10.10]{Davidson-Donsig-2010}, on a compact set,
multivariate polynomials uniformly approximate any continuous function.
Intuitively speaking, we use a multivariate polynomial to approximate, on the
unit ball of $\R^d$, some scaling of the indicator function corresponding to the intersection of
$m=K-1$ halfspaces $\bigcap_{j \in \cbr{1,\ldots,K} \setminus \cbr{i}} \cbr{x: \ip{w_i^* - w_j^*}{x} \geq 0}$. Within margin $\gamma$ along the decision boundary, we allow
the polynomial to attain arbitrary value. To be able to quantify $\gamma'$
we need to quantify certain parameters of the approximating polynomial.
Following the constructions in~\citet{Klivans-Servedio-2008}
(which in turn use the constructions in~\citet{Beigel-Reingold-Spielman-1995}), we
construct two different polynomials with different parameters that satisfy
Equation~\eqref{eqn:r-plus} and~\eqref{eqn:r-minus}.
The parameters
are quantified in
Theorems~\ref{theorem:polynomial-approximation-1}~and~\ref{theorem:polynomial-approximation-2}
stated below.

%The polynomial separates examples in
%one class from examples in the other classes.
%
%the indicator function corresponding to the intersection of
%$m=K-1$ halfspaces
%a function that takes value $\frac {\gamma'} 2$ on $R_i^+$
%and $-\frac {\gamma'} 2$ on $R_i^-$

\paragraph{Details of the proof.} Before we state the theorems, recall that a polynomial of $d$ variables is a
function $p:\R^d \to \R$ of the form
\begin{align*}
p(x)
& = p(x_1, x_2, \dots, x_d) \\
& = \sum_{\alpha_1, \alpha_2, \dots, \alpha_d} c_{\alpha_1, \alpha_2, \dots, \alpha_d} x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d}
\end{align*}
where the sum ranges over a finite set of $d$-tuples $(\alpha_1, \alpha_2,
\dots, \alpha_d)$ of non-negative integers and $c_{\alpha_1, \alpha_2, \dots,
\alpha_d}$'s are real coefficients. The \emph{degree} of a polynomial $p$, denoted
by $\deg(p)$, is the largest value of $\alpha_1 + \alpha_2 + \dots + \alpha_d$
for which the coefficient $c_{\alpha_1, \alpha_2, \dots, \alpha_d}$ is non-zero.
Following the terminology of~\citet{Klivans-Servedio-2008},
the \emph{norm of a polynomial} $p$ is defined as
$$
\norm{p} = \sqrt{\sum_{\alpha_1, \alpha_2, \dots, \alpha_d} \left(c_{\alpha_1, \alpha_2, \dots, \alpha_d} \right)^2 } \; .
$$
It is easy see that this is indeed a norm, since we can interpret it as the
Euclidean norm of the vector of the coefficients of the polynomial.

\begin{theorem}[Polynomial approximation of intersection of halfspaces I]
\label{theorem:polynomial-approximation-1}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$. There exists a
multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $p(x) \ge 1/2$ \\ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$,
\item $p(x) \le -1/2$ \\ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$,
\item $\displaystyle \deg(p) = \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil$,
\item $\displaystyle \norm{p} \le \left( 188 \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil \right)^{\frac{1}{2} \left\lceil \log_2(2m) \right\rceil \cdot \left\lceil \sqrt{\frac{1}{\gamma}} \right\rceil}$.
\end{enumerate}
\end{theorem}

\begin{theorem}[Polynomial approximation of intersection of halfspaces II]
\label{theorem:polynomial-approximation-2}
Let $v_1, v_2, \dots, v_m \in \R^d$ be vectors such that $\norm{v_1},
\norm{v_2}, \dots, \norm{v_m} \le 1$. Let $\gamma \in (0,1)$.
Define
$$
r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1 \quad \text{and} \quad s = \left \lceil \log_2(1/\gamma) \right \rceil \; .
$$
Then, there exists a multivariate polynomial $p:\R^d \to \R$ such that
\begin{enumerate}
\item $\displaystyle p(x) \ge 1/2$ \\
for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$,

\item $\displaystyle p(x) \le - 1/2$ \\
for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$,

\item $\deg(p) \le (2s+1) rm$,
\item $\norm{p} \le (4m-1) 2^m \cdot \left(2^s rm (4s+2)^2 \right)^{(s+1/2)rm}$.
\end{enumerate}
\end{theorem}

@@

The following lemma establishes a correspondence
between any multivariate polynomial in $\R^d$
and an element in $\ell_2$, and gives an upper bound on its norm.
Its proof follows from simple algebra, which we defer to
Appendix~\ref{section:proof-norm-bound}.

@@

Using the lemma and the polynomial approximation theorems, we can prove the
mapping $\phi$ maps any set of weakly linearly separable examples
to a strongly linearly separable set of examples. Due to space constraints,
we only sketch the
proof of Theorem~\ref{theorem:margin-transformation} here, and defer the full
proof to Appendix~\ref{section:full-proof-of-theorem-margin-transformation}.

\begin{proof}[Proof sketch of Theorem~\ref{theorem:margin-transformation}]
Since the examples $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ are weakly
linearly separable with margin $\gamma$,, there are vectors $w_1, w_2, \dots, w_K$
satisfying \eqref{equation:weak-linear-separability-1} and
\eqref{equation:weak-linear-separability-2}.

Fix any $i \in \{1,2,\dots,K\}$. Consider the $K-1$ vectors $(w_i - w_j)/2$ for
$j \in \{1,2,\dots,K\} \setminus \{i\}$. Note that the vectors have norm at most
$1$. We consider two cases regarding the relationship between $\gamma_1$ and
$\gamma_2$.

\paragraph{Case 1: $\gamma_1 \geq \gamma_2$.} In this setting, Theorem~\ref{theorem:polynomial-approximation-1}
implies that there exist a multivariate polynomial $p_i:\R^d \to \R$

\begin{align*}
\deg(p_i) & = \lceil \log_2(2K-2) \rceil \cdot \left\lceil \sqrt{\frac{2}{\gamma}} \right\rceil \; ,
\end{align*}
such that all examples $x$ in $R_i^+$ (resp. $R_i^-$) satisfy $p_i(x) \geq 1/2$
(resp. $p_i(x) \leq -1/2$).
Therefore, for all $t=1,2,\dots,T$, if $y_t = i$ then $p_i(x_t) \ge 1/2$,
 and if $y_t \neq i$ then $p_i(x_t) \le -1/2$, and
\begin{align*}
\norm{p_i} & \le \left(188 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; , \\
\end{align*}
By \autoref{lemma:norm-bound}, there exists $c_i \in \ell_2$ such that
$\ip{c_i}{\phi(x)} = p_i(x)$, and
\begin{align*}
\norm{c_i}_{\ell_2}
& \le \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil} \; .
\end{align*}
Define vectors $u_i \in \ell_2$ as
\begin{align*}
u_i & = \frac{c_i}{\sqrt{K} \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{\frac{1}{2} \lceil \log_2(2K-2) \rceil
\cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}} \; . \\
\end{align*}

Then, $\norm{u_1}^2 + \norm{u_2}^2 + \dots + \norm{u_K}^2 \le 1$.
Futhermore, for all $t=1,2,\dots,T$, $\ip{u_{y_t}}{x_t} \ge \gamma_1$
and for all $j \in \{1,2,\dots,K\} \setminus \{y_t\}$,
$\ip{u_j}{x_t} \le - \gamma_1$. In other words,
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_1 = \max\{\gamma_1, \gamma_2\}$.

\paragraph{Case 2: $\gamma_1 \leq \gamma_2$.} Following the
same idea as Case 1 and using Theorem~\ref{theorem:polynomial-approximation-1},
we can show that
$(\phi(x_1), y_1), (\phi(x_2), y_2), \dots, (\phi(x_T), y_T)$ are
strongly linearly separable with margin $\gamma_2 = \max\{\gamma_1, \gamma_2\}$.
%The proof for the case of $\gamma_1 < \gamma_2$
%is symmetric.

In summary, the examples are strongly
linearly separable with margin $\gamma' = \max\{\gamma_1, \gamma_2\}$.
Finally, observe that for any $t=1,2,\dots,T$,
\[
k(x_t,x_t) = \frac{1}{1 - \frac{1}{2} \norm{x_t}^2} \le 2 \; .
\qedhere
\]
\end{proof}
